WHITE WINE QUALITY ANALYSIS by Kasey Cox
========================================================

__Overview of data to be analyzed:__ This tidy data set contains 4,898 white 
wines with 11 variables on quantifying the chemical properties of each wine. At 
least 3 wine experts rated the quality of each wine, providing a rating between 
0 (very bad) and 10 (very excellent). This is a curated data set provided by
Udacity using the following research article:

Cortez, P.; Cerdeira, A.; Almeida, F.; Matos, T.; Reis, J.  
Modeling wine preferences by data mining from physicochemical properties.  
_Decision Support Systems_. __2009__, _47_, 547-553.

[Elsevier](http://dx.doi.org/10.1016/j.dss.2009.05.016)  
[Pre-press (pdf)](http://www3.dsi.uminho.pt/pcortez/winequality09.pdf)  
[bib](http://www3.dsi.uminho.pt/pcortez/dss09.bib)

__Question guiding investigation:__ Which chemical properties influence the 
quality of white wines?

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

library(ggplot2)
library(gridExtra)
library(plyr)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load the Data
white.wine.df <- read.csv("wineQualityWhites.csv")
```

***

# Univariate Plots Section

### Preliminary information on the data set  
```{r echo=FALSE, message=FALSE, warning=FALSE}
str(white.wine.df)
summary(white.wine.df)
```

### Organic acid contents and overall acidity (pH)  
```{r echo=FALSE, message=FALSE, warning=FALSE}
ta.hist <- ggplot(data = white.wine.df, aes(x = fixed.acidity)) +
  geom_histogram() +
  labs(x = "Fixed acidity (g/dm^3 of tartaric acid)")

ca.hist <- ggplot(data = white.wine.df, aes(x = citric.acid)) +
  geom_histogram() +
  labs(x = "Citric acid content (g/dm^3)")

aa.hist <- ggplot(data = white.wine.df, aes(x = volatile.acidity)) +
  geom_histogram() +
  labs(x = "Volatile Acidity (g/dm^3 of acetic acid)")

ph.hist <- ggplot(data = white.wine.df, aes(x = pH)) +
  geom_histogram() +
  labs(x = "pH")

grid.arrange(ta.hist, ca.hist, aa.hist, ph.hist, ncol =2)
```

The various acid measures all have a normal distribution, with all but pH having
a few outliers (in a visual sense).

__Citric acid:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}

print(paste("A few of the more extreme values...", head(subset(white.wine.df, 
            white.wine.df$citric.acid > 0.9, select = citric.acid))))

print(paste("3rd quartile:", quantile(white.wine.df$citric.acid, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$citric.acid, 0.99)))

```

The mean (0.3342) and median (0.3200) are similar in the citric acid 
distribution, meaning few wines exist in the tail and that most of the wines 
follow the normal distribution.

__Fixed acidity (tartaric acid):__  
```{r echo=FALSE, message=FALSE, warning=FALSE}
#subset(white.wine.df, 
       #white.wine.df$fixed.acidity > 10.0)[c("X", "fixed.acidity")]

print(paste("A few of the more extreme values...", head(subset(white.wine.df, 
            white.wine.df$fixed.acidity > 10.0, select = fixed.acidity))))

print(paste("3rd quartile:", quantile(white.wine.df$fixed.acidity, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$fixed.acidity, 0.99)))
```

The mean (6.855) and median (6.800) of fixed acidities are close in the 
distribution, meaning few wines exist in the tail and that most of the wines 
follow the normal distribution.

__Volatile acidity (acetic acid):__  
```{r echo=FALSE, message=FALSE, warning=FALSE}

print(paste("A few of the more extreme values...", head(subset(white.wine.df, 
            white.wine.df$volatile.acidity > 0.9, select = volatile.acidity))))

print(paste("3rd quartile:", quantile(white.wine.df$volatile.acidity, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$volatile.acidity, 0.99)))
```

The mean (0.2782) and median (0.2600) of volatile acidities are close in the 
distribution, meaning few wines exist in the tail and that most of the wines 
follow the normal distribution.

### Sulfur measures
```{r Sulfur measures, echo=FALSE, message=FALSE, warning=FALSE}
sulf.hist <- ggplot(data = white.wine.df, aes(x = sulphates)) +
  geom_histogram() +
  labs(x = "Sulphates (added potassium sulphate - g/dm^3)")

free.sulf.diox <- ggplot(data = white.wine.df, aes(x = free.sulfur.dioxide)) +
  geom_histogram() +
  labs(x = "free sulfur dioxide (mg / dm^3)")

total.sulf.diox <- ggplot(data = white.wine.df, aes(x = total.sulfur.dioxide)) +
  geom_histogram() +
  labs(x = "total sulfur dioxide (mg / dm^3)")

grid.arrange(sulf.hist, free.sulf.diox, total.sulf.diox, ncol = 2)
```

The sulfur measures also have normal distribution, though an extreme outlier
exists for the free sulfur dioxide histogram.

__Sulfates (added):__  
The distribution of sulfates is roughly normal.  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
summary(white.wine.df$sulphates)

print(paste("3rd quartile:", quantile(white.wine.df$sulphates, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$sulphates, 0.99)))
```


__Free sulfur dioxide:__  
The distribution of free sulfur dioxide is roughly normal, with an extreme
outlier (max).
```{r, echo=FALSE, message=FALSE, warning=FALSE}
summary(white.wine.df$free.sulfur.dioxide)

print(paste("3rd quartile:", quantile(white.wine.df$free.sulfur.dioxide, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$free.sulfur.dioxide, 0.99)))
```

__Total sulfur dioxide:__  
The distribution of total sulfur dioxide is roughly normal, with an extreme
outlier (max).
```{r, echo=FALSE, message=FALSE, warning=FALSE}
summary(white.wine.df$total.sulfur.dioxide)

print(paste("3rd quartile:", quantile(white.wine.df$total.sulfur.dioxide, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$total.sulfur.dioxide, 0.99)))
```


### Other wine attributes
__Density and residual sugar:__  
```{r echo=FALSE, message=FALSE, warning=FALSE}
res.sugar.hist <- ggplot(data = white.wine.df, aes(x = residual.sugar)) +
  geom_histogram() +
  labs(x = "Residual sugars (g/dm^3)")

density.hist <- ggplot(data = white.wine.df, aes(x = density)) +
  geom_histogram() +
  labs(x = "density (g/cm^3)")

grid.arrange(res.sugar.hist, density.hist, ncol = 2)
```

In the distributions for both density and residual sugar, we can see an extreme 
outlier (maybe more than one) in both. The distributions are otherwise 
relatively narrow. My suspicion is that they are the same wine (or few wines). 
Dissolved sugar increases the density of water (which is 1.00 g/cm^3).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(paste("High density wine:", 
            white.wine.df$density[white.wine.df$density > 1.02]))

print(paste("High residual sugar wine:", 
            white.wine.df$residual.sugar[white.wine.df$residual.sugar > 60]))

subset(white.wine.df, white.wine.df$density > 1.02)[c("residual.sugar", 
                                                      "density")]
```

We can see that they are indeed the same wine.

__Sodium chloride:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
nacl.hist <- ggplot(data = white.wine.df, aes(x = chlorides)) +
  geom_histogram() +
  labs(x = "Sodium chloride (g/dm^3)")

summary(white.wine.df$chlorides)
```

There is a little bit of a tail in the sodium chloride data; most of the 
values are less than 0.05, yet there are a handful of values greater than 0.2 
Let's do a log base 10 transform on the x-axis.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
nacl.log10.hist <- ggplot(data = white.wine.df, aes(x = chlorides)) +
  geom_histogram() +
  labs(x = "Log 10 Sodium Chloride") +
  scale_x_log10()

grid.arrange(nacl.hist, nacl.log10.hist, ncol = 1)
```

We see that there really aren't many data points beyond 0.1 g/dm^3. This is
consistent with the mean not being far from the median.

__Alcohol (ABV - % by volume):__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = alcohol)) +
  geom_histogram(binwidth = 0.25) +
  labs(x = "Alcohol (ABV - % by volume)")
```

The distribution of ABV is roughly normal.

__Quality Ratings:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Quality Rating (0 - 10)") +
  scale_x_continuous(limits = c(2, 10), breaks = seq(0, 10, 1))
```

The quality ratings are uniformly distributed around 6 (the mode).

### Comparing the best, middle, and worst wines
Wine ratings were made on a scale of 0 to 10, with 0 being the worst. In this
data, wines ranged from 3 to 9. I divided the ratings as follows:  
* Worst: wines rated 3, 4  
* Middle: wines rated 5, 6, 7  
* Best: wines rated 8, 9  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Make quality categorical
white.wine.df$quality <- factor(x = white.wine.df$quality, levels = c(0:10), ordered = TRUE)

# Add new categorical variable
white.wine.df$quality.category <- ifelse(white.wine.df$quality >= 8, "Best", 
                                         ifelse(white.wine.df$quality >= 5, 
                                                "Middle", 
                                                ifelse(white.wine.df$quality 
                                                       >= 3, "Worst", "Other")))

# Number of wines in each category (need plyr loaded)
print(paste("Best rated wines:", 
            count(white.wine.df$quality.category == "Best")))
print(paste("Middle rated wines:", 
            count(white.wine.df$quality.category == "Middle")))
print(paste("Worst rated wines:", 
            count(white.wine.df$quality.category == "Worst")))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Boxplot theme
boxplot.theme <- theme(panel.background = element_rect(fill="gray90", color="gray90"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color = "gray50"),
        panel.grid.minor.x = element_blank(), 
        plot.title = element_text(size=18, hjust=0.5),
        legend.title = element_blank())

boxplot.colors <- scale_fill_manual(values=c("green3", "deepskyblue3", "red2"))
```

__Volatile acidity:__  
In the text accompanying the data set, there is information on the variables
contained in the data. Of note is this: "volatile acidity: the amount of acetic 
acid in wine, which at too high of levels can lead to an unpleasant, vinegar 
taste."

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, y = volatile.acidity)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "Volatile Acidity (acetic acid; g/dm^3)",
       title = "Volatile Acidity by Quality") +
  boxplot.colors +
  boxplot.theme +
  stat_summary(fun.y = mean, geom = "point", shape = 4)
```

It seems that poorly rated wines tend to have higher volatile acidity. This is 
something that will be further explored in the bivariate analysis.

__pH:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, y = pH)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "pH", title = "pH by Quality") +
  boxplot.colors +
  boxplot.theme +
  stat_summary(fun.y = mean, geom = "point", shape = 4)
```

Better wines appear to have a (relatively) higher pH, with the middle rated
wines having a lot of variability. This is something that will be further 
explored in the bivariate analysis.

__Free sulfur dioxide:__  
Another note from the text accompanying the data set: "total sulfur dioxide: 
amount of free and bound forms of S02; in low  concentrations, SO2 is mostly 
undetectable in wine, but at free SO2 concentrations over 50 ppm, SO2 becomes 
evident in the nose and taste of wine"

At the cited 50 ppm, that would mean free sulfur dioxide would have to be at
least 49.9 mg/dm^3 to be detectable.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, 
                                 y = free.sulfur.dioxide)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "Free Sulfur Dioxide", 
       title = "Free Sulfur Dioxide by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

Somewhat counterintuitively, the median and mean free sulfur dioxide is higher
in the better rated wines; however there is more variability in the worst
wines. Note that the wine with the highest free sulfur dioxide is in the
worst category.

__Residual sugar:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, y = residual.sugar)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip(ylim = c(0, 30)) +
  labs(x = "Quality Category", y = "Residual sugar (g/dm^3)", 
       title = "Residual Sugar by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

There doesn't appear to be a correlation at first glance.

__Alcohol (ABV - % by volume):__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, y = alcohol)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "Alcohol (ABV)", 
       title = "ABV by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

The better rated wines seem to generally have higher alcohol content.

__Chlorides:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, y = chlorides)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip(ylim = c(0, 0.15)) +
  labs(x = "Quality Category", y = "Sodium chloride (g / dm^3)", 
       title = "Sodium Chloride Content by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

The worse rated wines tend to have higher sodium chloride content. I found
some information to corroborate this observation:  
[Peer reviewed paper](http://www.scielo.br/scielo.php?script=sci_arttext&pid=S0101-20612015000100095)  
[Blog post](http://wineoscope.com/2015/10/02/when-a-wine-is-salty-and-why-it-shouldnt-be/)


# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?
This data has 4898 observations with 13 features (variables). These features are
"X", "fixed.acidity", "volatile.acidity", "citric.acid", "residual.sugar",
"chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide" "density", "pH", 
"sulphates", "alcohol", and "quality".

Most of the variables are numerical, except for "quality". Although it consists
of numbers, they represent rank rather than being numbers in the strict sense.

### What is/are the main feature(s) of interest in your dataset?


### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!