WHITE WINE QUALITY ANALYSIS by Kasey Cox
========================================================

__Overview of data to be analyzed:__ This tidy data set contains 4,898 white 
wines with 11 variables on quantifying the chemical properties of each wine. At 
least 3 wine experts rated the quality of each wine, providing a rating between 
0 (very bad) and 10 (very excellent). This is a curated data set provided by
Udacity using the following research article:

Cortez, P.; Cerdeira, A.; Almeida, F.; Matos, T.; Reis, J.  
Modeling wine preferences by data mining from physicochemical properties.  
_Decision Support Systems_. __2009__, _47_, 547-553.

[Elsevier](http://dx.doi.org/10.1016/j.dss.2009.05.016)  
[Pre-press (pdf)](http://www3.dsi.uminho.pt/pcortez/winequality09.pdf)  
[bib](http://www3.dsi.uminho.pt/pcortez/dss09.bib)

__Question guiding investigation:__ Which chemical properties influence the 
quality of white wines?

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

library(ggplot2)
library(gridExtra)
library(plyr)
library(GGally)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load the Data
white.wine.df <- read.csv("wineQualityWhites.csv")

# Remove "X" variable
white.wine.df$X <- NULL
```

***

# Univariate Plots Section

### Preliminary information on the data set  
```{r echo=FALSE, message=FALSE, warning=FALSE}
str(white.wine.df)
summary(white.wine.df)
```

### Organic acid contents and overall acidity (pH)  
```{r echo=FALSE, message=FALSE, warning=FALSE}
ta.hist <- ggplot(data = white.wine.df, aes(x = fixed.acidity)) +
  geom_histogram() +
  labs(x = expression("Fixed acidity in"~g/dm^3~"tartaric acid"))

ca.hist <- ggplot(data = white.wine.df, aes(x = citric.acid)) +
  geom_histogram() +
  labs(x = expression("Citric Acid in"~g/dm^3))

aa.hist <- ggplot(data = white.wine.df, aes(x = volatile.acidity)) +
  geom_histogram() +
  labs(x = expression("Volatile acidity in"~g/dm^3~"acetic acid"))

ph.hist <- ggplot(data = white.wine.df, aes(x = pH)) +
  geom_histogram() +
  labs(x = "pH")

grid.arrange(ta.hist, ca.hist, aa.hist, ph.hist, ncol =2)
```

The various acid measures all have a normal distribution, with all but `pH` 
having a few outliers (in a visual sense).

__Citric acid:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}

cat("A few of the more extreme values...", 
    as.character(head(subset(white.wine.df, 
            white.wine.df$citric.acid > 0.9, select = citric.acid))), 
    "\n3rd quartile:", as.character(quantile(white.wine.df$citric.acid, 0.75)), 
    "\n99th percentile:", 
            as.character(quantile(white.wine.df$citric.acid, 0.99)))
```

The mean (0.3342) and median (0.3200) are similar in the `citric.acid` 
distribution, meaning few wines exist in the tail and that most of the wines 
follow the normal distribution.

__Fixed acidity (tartaric acid):__  
```{r echo=FALSE, message=FALSE, warning=FALSE}
cat("A few of the more extreme values...", 
    as.character(head(subset(white.wine.df, white.wine.df$fixed.acidity > 10.0, 
                             select = fixed.acidity))), 
    "\n3rd quartile:", 
    as.character(quantile(white.wine.df$fixed.acidity, 0.75)), 
    "\n99th percentile:", 
            as.character(quantile(white.wine.df$fixed.acidity, 0.99)))
```

The mean (6.855) and median (6.800) of `fixed.acidity` are close in the 
distribution, meaning few wines exist in the tail and that most of the wines 
follow the normal distribution.

__Volatile acidity (acetic acid):__  
```{r echo=FALSE, message=FALSE, warning=FALSE}

cat("A few of the more extreme values...", 
    as.character(head(subset(white.wine.df, 
            white.wine.df$volatile.acidity > 0.9, select = volatile.acidity))), 
    "\n3rd quartile:", 
    as.character(quantile(white.wine.df$volatile.acidity, 0.75)), 
    "\n99th percentile:", 
            as.character(quantile(white.wine.df$volatile.acidity, 0.99)))
```

The mean (0.2782) and median (0.2600) of `volatile.acidity` are close in the 
distribution, meaning few wines exist in the tail and that most of the wines 
follow the normal distribution.

### Sulfur measures
```{r Sulfur measures, echo=FALSE, message=FALSE, warning=FALSE}
sulf.hist <- ggplot(data = white.wine.df, aes(x = sulphates)) +
  geom_histogram() +
  labs(x = expression("Sulphates in"~g/dm^3))

free.sulf.diox <- ggplot(data = white.wine.df, aes(x = free.sulfur.dioxide)) +
  geom_histogram() +
  labs(x = expression("Free sulfur dioxide in"~mg/dm^3))

total.sulf.diox <- ggplot(data = white.wine.df, aes(x = total.sulfur.dioxide)) +
  geom_histogram() +
  labs(x = expression("Total sulfur dioxide in"~mg/dm^3))

grid.arrange(sulf.hist, free.sulf.diox, total.sulf.diox, ncol = 2)
```

The sulfur measures also have normal distribution, though an extreme outlier
exists for the `free.sulfur.dioxide` histogram.

__Sulfates (added):__  
The distribution of `sulfates` is roughly normal.  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
summary(white.wine.df$sulphates)

cat("\n3rd quartile:", as.character(quantile(white.wine.df$sulphates, 0.75)), 
    "\n99th percentile:", 
    as.character(quantile(white.wine.df$sulphates, 0.99)))
```

__Free sulfur dioxide:__  
The distribution of `free.sulfur.dioxide` is roughly normal, with an extreme
outlier (max).
```{r, echo=FALSE, message=FALSE, warning=FALSE}
summary(white.wine.df$free.sulfur.dioxide)

cat("\n3rd quartile:", 
    as.character(quantile(white.wine.df$free.sulfur.dioxide, 0.75)), 
    "\n99th percentile:", 
            as.character(quantile(white.wine.df$free.sulfur.dioxide, 0.99)))
```

__Total sulfur dioxide:__  
The distribution of `total.sulfur.dioxide` is roughly normal, with an extreme
outlier (max).
```{r, echo=FALSE, message=FALSE, warning=FALSE}
summary(white.wine.df$total.sulfur.dioxide)

cat("\n3rd quartile:", 
    as.character(quantile(white.wine.df$total.sulfur.dioxide, 0.75)), 
    "\n99th percentile:", 
            as.character(quantile(white.wine.df$total.sulfur.dioxide, 0.99)))
```


### Other wine attributes
__Density and residual sugar:__  
```{r echo=FALSE, message=FALSE, warning=FALSE}
res.sugar.hist <- ggplot(data = white.wine.df, aes(x = residual.sugar)) +
  geom_histogram() +
  labs(x = expression("Residual sugars in"~g/dm^3))

density.hist <- ggplot(data = white.wine.df, aes(x = density)) +
  geom_histogram() +
  labs(x = expression("Density in"~g/cm^3))

grid.arrange(res.sugar.hist, density.hist, ncol = 2)
```

In the distributions for both `density` and `residual.sugar`, we can see an 
extreme outlier (maybe more than one) in both. The distributions are otherwise 
relatively narrow. My suspicion is that they are the same wine (or few wines). 
Dissolved sugar increases the density of water (which is 1.00 g/cm^3).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("High density wine:", 
            as.character(white.wine.df$density[white.wine.df$density > 1.02]), 
    "\nHigh residual sugar wine:", 
            as.character(white.wine.df$residual.sugar[white.wine.df$residual.sugar > 60]))

subset(white.wine.df, white.wine.df$density > 1.02)[c("residual.sugar", 
                                                      "density")]
```

We can see that they are indeed the same wine.

__Sodium chloride:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
nacl.hist <- ggplot(data = white.wine.df, aes(x = chlorides)) +
  geom_histogram() +
  labs(x = expression("Sodium chloride in"~g/dm^3))

summary(white.wine.df$chlorides)
```

There is a little bit of a tail in the `chlorides` distribution; most of the 
values are less than 0.05, yet there are a handful of values greater than 0.2 
Let's do a log base 10 transform on the x-axis.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
nacl.log10.hist <- ggplot(data = white.wine.df, aes(x = chlorides)) +
  geom_histogram() +
  labs(x = "Log base 10 Sodium Chloride") +
  scale_x_log10()

grid.arrange(nacl.hist, nacl.log10.hist, ncol = 1)
```

We see that there really aren't many data points beyond 0.1. This is
consistent with the mean not being far from the median.

__Alcohol (ABV - % by volume):__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = alcohol)) +
  geom_histogram(binwidth = 0.25) +
  labs(x = "Alcohol\n(ABV % by volume)")
```

The distribution of `alcohol` is roughly normal.

__Quality Ratings:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Quality Rating (0 - 10)") +
  scale_x_continuous(limits = c(2, 10), breaks = seq(0, 10, 1))
```

`quality` ratings are uniformly distributed around 6 (the mode).

### More about the best, middle, and worst wines
Wine ratings were made on a scale of 0 to 10, with 0 being the worst. In this
data, wines ranged from 3 to 9. I divided the ratings as follows:  
> Worst: wines rated 3, 4  
> Middle: wines rated 5, 6, 7  
> Best: wines rated 8, 9  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Make quality categorical
white.wine.df$quality <- factor(x = white.wine.df$quality, levels = c(0:10), 
                                ordered = TRUE)

# Add new categorical variable
white.wine.df$quality.category <- ifelse(white.wine.df$quality >= 8, "Best", 
                                         ifelse(white.wine.df$quality >= 5, 
                                                "Middle", 
                                                ifelse(white.wine.df$quality 
                                                       >= 3, "Worst", "Other")))

# Number of wines in each category (need plyr loaded)
cat("Best rated wines:", 
            as.character(count(white.wine.df$quality.category == "Best")), 
    "\nMiddle rated wines:", 
            as.character(count(white.wine.df$quality.category == "Middle")), 
    "\nWorst rated wines:", 
            as.character(count(white.wine.df$quality.category == "Worst")))
```

I am going to focus on some of the features of the best wines.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
best.va.hist <- ggplot(data = subset(white.wine.df, quality.category == "Best"), 
                       aes(x = volatile.acidity)) + 
  geom_histogram(bins = 40)

best.ph.hist <- ggplot(data = subset(white.wine.df, quality.category == "Best"), 
                       aes(x = pH)) +
  geom_histogram(bins = 40)

best.fsd.hist <- ggplot(data = subset(white.wine.df, quality.category == "Best"), 
                        aes(x = free.sulfur.dioxide)) +
  geom_histogram(bins = 40)

best.rs.hist <- ggplot(data = subset(white.wine.df, quality.category == "Best"), 
                       aes(x = residual.sugar)) +
  geom_histogram(bins = 40)

best.alcohol.hist <- ggplot(data = subset(white.wine.df, 
                                          quality.category == "Best"), 
                            aes(x = alcohol)) +
  geom_histogram(bins = 40)

best.nacl.hist <- ggplot(data = subset(white.wine.df, quality.category == "Best"), 
                         aes(x = chlorides)) +
  geom_histogram(bins = 40)

grid.arrange(best.va.hist, best.ph.hist, best.fsd.hist, best.rs.hist, best.alcohol.hist, best.nacl.hist, ncol = 2)
```

It appears that in the best wines (rated 8 or greater), certain features show
less variance whereas others vary greatly. For example, `chlorides` does 
not seem to vary much (except for an extreme outlier); `residual.sugar` has a 
lot of variation.

Most of the distribution show a normal or somewhat normal distribution. The
`chlorides` distribution is relatively normal. The `residual.sugar` 
distribution seems almost uniform in its distribution.

# Univariate Analysis

### What is the structure of your dataset?
This data has 4898 observations (wines) with 13 features (variables). These 
features are "X", "fixed.acidity", "volatile.acidity", "citric.acid", 
"residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", 
"density", "pH", "sulphates", "alcohol", and "quality".

Most of the variables are numerical, except for "quality". Although it consists
of numbers, the numbers represent rank rather than being numbers in the strict 
mathematical sense. When I loaded the data, R interpreted it as a numerical 
vector, so I purposely changed it to a factor to properly reflect what it 
stands for.

### What is/are the main feature(s) of interest in your dataset?

The main feature of my data set is "quality" which, as mentioned in the previous
answer, represets a ranking: 0 to 10 with 0 being the worst quality rating and 
10 being the best quality rating.

The goal of this analysis is to determine which chemical properties affect the
wine's rating (quality).

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Because we want to know which chemical properties affect quality, a feature's 
importance will be determined relative to its effect on quality.

So far, a few features stand out: alchohol content, chloride content, free 
sulfur dioxide, and volatile acidity.

### Did you create any new variables from existing variables in the dataset?

Yes. I created a "quality.category" (factor) variable using the "quality" 
variable. This factor has three levels: "Worst", "Middle", and "Best". If a 
wine was rated poorly (0-3), its quality.category value is "Worst". A rating 
of 5-7 corresponds to "Middle". A rating of 8-10 represents "Best".

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

The histogram of chloride content had a long tail with what appeared to be many 
data points along its length. I did a log transformation along its x-axis to 
see if there really were a good amount of data points in the tail. After the 
transformation, I found that the tail was less significant than I had 
originally thought.

***

# Bivariate Plots Section

### Preliminary plotting
For my preliminary bivariate plotting, I am using the `ggpairs` function to 
help guide my later, targeted plots. The resulting plot matrix will give 
me an initial look at any possible relationships that may be worth exploring 
further.

To prevent overcrowded plot matrices, I will be subsetting the data into 
different groups based on my hunches about variables that may have something 
to do with one another. I will also create a subset that includes variables
that are, at least to my understanding, seemingly unrelated.

__Acidity subset__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
acid.subsetted.data <- subset(white.wine.df, !is.na(white.wine.df), 
                                 select = c(fixed.acidity, volatile.acidity, 
                                            citric.acid, pH, quality))

ggpairs(acid.subsetted.data, 
        lower = list(continuous = wrap("points")))
```

There only appears to be one moderate correlation amongst the numeric variables 
in this subset: between `pH` and `fixed.acidity` (r = -0.426). 

There also only appears to be one correlation between `quality` (categorical 
variable) and `pH`. This will require further exploration below.

__Sulfur subset__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
sulfur.subsetted.data <- subset(white.wine.df, !is.na(white.wine.df), 
                                 select = c(free.sulfur.dioxide, 
                                            total.sulfur.dioxide, 
                                            sulphates, quality))

ggpairs(sulfur.subsetted.data, 
        lower = list(continuous = wrap("points")))
```

Amongst the numerical variables, a correlation exists between 
`free.sulfur.dioxide` and `total.sulfur.dioxide` (r = 0.616). 

For `quality`, the only relationship that appears to exist is between `quality` 
and `total.sulfur.dioxide`, but this requires more investigation below.

__Other features subset__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
other.subsetted.data <- subset(white.wine.df, !is.na(white.wine.df), 
                                 select = c(residual.sugar, chlorides, 
                                            density, alcohol, quality))

ggpairs(other.subsetted.data, 
        lower = list(continuous = wrap("points")))
```

A strong positive correlation exists between `residual.sugar` and `density` 
(r = 0.839). There is also a strong negative correlation between `alcohol` 
and `density` (r = -0.78).  Finally, there is a modest negative correlation 
between `alcohol` and `residual.sugar` (r = -0.4506).

As for relationships involving quality, there definitely appears to be a 
positive relationship between `quality` and `alcohol`. There also appears to 
be negative relationships between `quality` and `chlorides` as well as 
`quality` and `density`. A relationship between `quality` and `residual.sugar` 
seems unclear. All of these will be explored further below.

__Random features subset__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
random.subsetted.data <- subset(white.wine.df, !is.na(white.wine.df), 
                                 select = c(sulphates, pH, 
                                            total.sulfur.dioxide, 
                                            density, citric.acid, quality))

ggpairs(random.subsetted.data, 
        lower = list(continuous = wrap("points")))
```

A modest positive correlation seems to exists between `density` and 
`total.sulfur.dioxide` (r = 0.53).

No other significant correlations seems to exist amongst the numerical 
variables.

No new relationships involving `quality` are obvious.

### Relationships amongst chemical properties
Taking a close look at above findings between different pairs of chemical 
properties (based on correlation coefficient findings).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
scatterplot.theme <- theme(panel.background = element_rect(fill="white", 
                                                       color="white"),
        panel.grid.major.y = element_line(color = "gray70"),
        panel.grid.major.x = element_line(color = "gray70"),
        panel.grid.minor.x = element_blank(), 
        plot.title = element_text(size=18, hjust=0.5),
        legend.title = element_blank())
```

__pH and Fixed Acidity__  
[With 4 "far out" outliers eliminated](https://en.wikipedia.org/wiki/Outlier#Tukey.27s_test)  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
fa.iqr <- quantile(white.wine.df$fixed.acidity, 0.75) - quantile(white.wine.df$fixed.acidity , 0.25)
fa.no.outliers <- subset(white.wine.df, white.wine.df$fixed.acidity < 
                       (quantile(white.wine.df$fixed.acidity, 0.75) + 3*fa.iqr))

ggplot(data = fa.no.outliers, aes(x = fixed.acidity, y = pH)) +
  geom_point(color = "dodgerblue4", alpha = 0.2, position = "jitter") +
  geom_smooth(method = "lm", color = "navy") +
  labs(x = "Fixed acidity (g / dm^3)", y = "pH", title = "pH by Fixed Acidity") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("r =", as.character(cor(x = white.wine.df$fixed.acidity, 
                       y = white.wine.df$pH, method = "pearson")))
```

__Total sulfur dioxide and free sulfur dioxide__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide)) +
  geom_point(color = "dodgerblue4", alpha = 0.1, position = "jitter") +
  geom_smooth(method = "lm", color = "navy") +
  coord_cartesian(xlim = c(0,150)) +
  labs(x = "Free Sulfur Dioxide\n(mg / dm^3)", y = "Total Sulfur Dioxide\n(mg / dm^3)", title = "Total vs. Free Sulfur Dioxide") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("r =", as.character(cor(x = white.wine.df$free.sulfur.dioxide, 
                       y = white.wine.df$total.sulfur.dioxide, 
                       method = "pearson")))
```

__Density vs. residual sugar__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = residual.sugar, y = density)) +
  geom_point(color = "dodgerblue4", alpha = 0.1, position = "jitter") +
  geom_smooth(method = "lm", color = "navy") +
  coord_cartesian(xlim = c(0, 35), ylim = c(0.985, 1.015)) +
  labs(x = "Residual Sugar\n(g / dm^3)", y = "Density\n(g / cm^3)", 
       title = "Density vs. Residual Sugar") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("r =", as.character(cor(x = white.wine.df$residual.sugar, 
                       y = white.wine.df$density, 
                       method = "pearson")))
```

__Alcohol vs. residual sugar__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = residual.sugar, y = alcohol)) +
  geom_point(color = "dodgerblue4", alpha = 0.2, position = "jitter") +
  geom_smooth(method = "lm", color = "navy", se = FALSE, 
              formula = y ~ poly(x, 2)) +
  coord_cartesian(x = c(0, 35)) +
  labs(x = "Residual Sugar\n(g / dm^3)", y = "Alcohol\n(% by Volume)", 
       title = "Alcohol vs. Residual Sugar") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("r =", as.character(cor(x = white.wine.df$residual.sugar, 
                       y = white.wine.df$alcohol, 
                       method = "pearson")))
```

__Density vs. alcohol__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = alcohol, y = density)) +
  geom_point(color = "dodgerblue4", alpha = 0.1, position = "jitter") +
  geom_smooth(method = "lm", color = "navy") + 
  coord_cartesian(ylim = c(0.985, 1.015)) + 
  labs(x = "Alcohol\n(% by volume)", y = "Density\n(g / cm^3)", 
       title = "Density vs. Alcohol") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("r =", as.character(cor(x = white.wine.df$alcohol, 
                       y = white.wine.df$density, 
                       method = "pearson")))
```

__Density and Total Sulfur Dioxide__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = total.sulfur.dioxide, y = density)) +
  geom_point(color = "dodgerblue4", alpha = 0.1, position = "jitter") +
  geom_smooth(method = "lm", color = "navy") + 
  coord_cartesian(ylim = c(0.985, 1.015)) + 
  labs(x = "Total Sulfur Dioxide\n(mg / dm^3)", y = "Density\n(g / cm^3)", 
       title = "Density vs. Total Sulfur Dioxide") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("r =", as.character(cor(x = white.wine.df$total.sulfur.dioxide, 
                       y = white.wine.df$density, 
                       method = "pearson")))
```


### Relationships between quality and other chemical properties
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Boxplot theme
boxplot.theme <- theme(panel.background = element_rect(fill="gray90", 
                                                       color="gray90"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color = "gray50"),
        panel.grid.minor.x = element_blank(), 
        plot.title = element_text(size=18, hjust=0.5),
        legend.title = element_blank())

boxplot.colors <- scale_fill_manual(values=c("green3", "deepskyblue3", "red2"))
```

__Volatile acidity:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality, y = volatile.acidity)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "Volatile Acidity\n(g/dm^3)",
       title = "Volatile Acidity by Quality") +
  boxplot.colors +
  boxplot.theme +
  stat_summary(fun.y = mean, geom = "point", shape = 4)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median volatile acidity for 'Best' wines:", 
    as.character(median(subset(white.wine.df,
                               white.wine.df$quality.category == 
                               "Best")$volatile.acidity)), 
    "\nMedian volatile acidity for 'Middle' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Middle")$volatile.acidity)), 
    "\nMedian volatile acidity for 'Worst' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Worst")$volatile.acidity)))
```

It seems that poorly rated wines tend to have higher `volatile.acidity`.

__pH:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality, y = pH)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "pH", title = "pH by Quality") +
  boxplot.colors +
  boxplot.theme +
  stat_summary(fun.y = mean, geom = "point", shape = 4)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median pH for 'Best' wines:", 
    as.character(median(subset(white.wine.df,
                               white.wine.df$quality.category == 
                               "Best")$pH)), 
    "\nMedian pH for 'Middle' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Middle")$pH)), 
    "\nMedian pH for 'Worst' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Worst")$pH)))
```

The best wines appear to have a (relatively) higher `pH`; the middle rated
wines have a lot of variability.

__Free sulfur dioxide:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality, 
                                 y = free.sulfur.dioxide)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip(ylim = c(0, 150)) +
  labs(x = "Quality Category", y = "Free Sulfur Dioxide\n(mg / dm^3)", 
       title = "Free Sulfur Dioxide by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median Free Sulfur Dioxide for 'Best' wines:", 
    as.character(median(subset(white.wine.df,
                               white.wine.df$quality.category == 
                               "Best")$free.sulfur.dioxide)), 
    "\nMedian Free Sulfur Dioxide for 'Middle' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Middle")$free.sulfur.dioxide)), 
    "\nMedian Free Sulfur Dioxide for 'Worst' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Worst")$free.sulfur.dioxide)))

subset(white.wine.df, 
       white.wine.df$free.sulfur.dioxide > 150)[c("free.sulfur.dioxide", 
                                                  "quality", 
                                                  "quality.category")]
```

The median `free.sulfur.dioxide` is lower in the poorly rated wines when 
considered as a group (with respect to `quality.category`); however, 
there is a lot variability amongst these wines. Overall, when considering each 
`quality` rating separately, though, the pattern seems dubious. Note that the 
wine with the highest `free.sulfur.dioxide` (289) is in the worst category. 
This may distort the "Worst" category as a group.

__Total sulfur dioxide:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality, 
                                 y = total.sulfur.dioxide)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "Total Sulfur Dioxide\n(mg / dm^3)", 
       title = "Total Sulfur Dioxide by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median Total Sulfur Dioxide for 'Best' wines:", 
    as.character(median(subset(white.wine.df,
                               white.wine.df$quality.category == 
                               "Best")$free.sulfur.dioxide)), 
    "\nMedian Total Sulfur Dioxide for 'Middle' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Middle")$free.sulfur.dioxide)), 
    "\nMedian Total Sulfur Dioxide for 'Worst' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Worst")$free.sulfur.dioxide)))
```

There appears to be a slight positive trend between `total.sulfur.dioxide` and 
`quality`.

__Alcohol (ABV - % by volume):__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality, y = alcohol)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "Alcohol\n(ABV)", 
       title = "ABV by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median Alcohol for 'Best' wines:", 
    as.character(median(subset(white.wine.df,
                               white.wine.df$quality.category == 
                               "Best")$alcohol)), 
    "\nMedian Alcohol for 'Middle' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Middle")$alcohol)), 
    "\nMedian Alcohol for 'Worst' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Worst")$alcohol)))
```

The better rated wines generally have higher alcohol content.

__Density:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality, y = density)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip(ylim = c(0.985, 1.015)) +
  labs(x = "Quality Category", y = "Density\n(g/cm^3)", 
       title = "Density by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median Density for 'Best' wines:", 
    as.character(median(subset(white.wine.df,
                               white.wine.df$quality.category == 
                               "Best")$density)), 
    "\nMedian Density for 'Middle' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Middle")$density)), 
    "\nMedian Density for 'Worst' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Worst")$density)))
```

Better rated wines tend to have lower `density`.

__Chlorides:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality, y = chlorides)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip(ylim = c(0, 0.15)) +
  labs(x = "Quality Category", y = "Sodium chloride\n(g / dm^3)", 
       title = "Sodium Chloride Content by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median Chlorides for 'Best' wines:", 
    as.character(median(subset(white.wine.df,
                               white.wine.df$quality.category == 
                               "Best")$chlorides)), 
    "\nMedian Chlorides for 'Middle' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Middle")$chlorides)), 
    "\nMedian Chlorides for 'Worst' wines:", 
    as.character(median(subset(white.wine.df, 
                               white.wine.df$quality.category == 
                                 "Worst")$chlorides)))
```

The worse rated wines tend to have higher `chlorides`.

***

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

I observed several relationships involving the main feature (`quality`). Poorly 
rated wines ("Worst" `quality.category`) tended to have higher 
`volatile.acidity`. In the text accompanying the data set, there is 
information on the variables contained in the data. Of note is this: "volatile 
acidity: the amount of acetic acid in wine, which at too high of levels can 
lead to an unpleasant, vinegar taste." This would explain why at higher values 
these wines would end up in the worst category. Perhaps "Middle" and "Best" 
wines do not hit the threshold of "too high of levels", thereby not 
experiencing any negative impact from `volatile.acidity`.

Wines with a higher `pH` value tended to have a better `quality` rating. This 
holds true for all three `quality.category` subsets ("Worst", "Middle", 
"Best"). I do _not_ believe this is because wines with higher 
`volatile.acidity` generally have worse ratings. Below (next question) I go 
into a discussion about how `volatile.acidity` has a low impact on final `pH` 
(how, in fact, it is `fixed.acidity` that largely determines final `pH`). 
Maybe it is a tasting 
preference to have slightly less acidic wine (they are all acidic 
nonetheless).

No relationship appears to exist between `quality` and 
`free.sulfur.dioxide`. Another note from the text accompanying the data set 
states: "total sulfur dioxide: amount of free and bound forms of S02; in low 
concentrations, SO[2] is mostly undetectable in wine, but at free SO[2] 
concentrations over 50 ppm, SO[2] becomes evident in the nose and taste 
of wine"

Somewhat counterintuitively, there doesn't seem to be much of a trend. A 
closer examination reveals that at the cited 50 ppm, that would mean free 
sulfur dioxide would have to be at least 49.9 mg/dm^3 to be detectable. Most 
wines (for all ratings) are less than this threshold. So for the most part, 
too much free SO2 likely isn't a factor unless it is at unreasonably high 
amounts.

For `quality` and `total.sulfur.dioxide`, however, there does appear to be a 
modest trend. I did a little searching to make sense of this observation. 
Sulfur dioxide is an important _preservative_ and _stabilizer_ used world-wide 
in wine making. `total.sulfur.dioxide` is defined as the "amount of free and 
bound forms of S02".

In cases where the wine is highly oxidized and has ample amounts of 
acetaldehyde (not good for wine quality), any added SO2 is quickly bound by 
acetaldehyde and other chemical constituents. It becomes part of total SO2 
but not part of free SO2. Bound SO2 does _not_ have antioxidant or 
anti-microbial properties, but free SO2 does. Poorly made wines that 
have oxidized require more SO2 to be added -- enough so that it spills 
over into free SO2 at a level high enough to have the desired preserving 
effect. Thereby explaining why we see a modest trend for 
`total.sulfur.dioxide` but not for `free.sulfur.dioxide`.

`quality` ratings tend to be higher when `alcohol` is higher. This appears to 
be a preference. According to an article in the Scientific American (
"Wine Becomes More Like Whisky as Alcohol Content Gets High"):  
> "The recent surge in wine's punch [alcohol content] is largely a result, 
scientists say, of a fashion for deeply colored wines with fewer “green” 
qualities and more bright, ripe, fruity flavors. As New World wines in this 
style have drawn more fans, even European winemakers accustomed to making 
lower-alcohol wines in less ripe styles are beginning to follow suit. But 
producing wines with those flavors means letting grapes hang longer on the 
vine, and with longer hang times comes bigger sugar. The more sugar the wine 
yeast _S. cerevisiae_ has to work with, the more alcohol it will make."

Along similiar lines `quality` tends to be higher in wines with lower 
`density`. This is consistent with the previous relationship that showed 
wines with higher alcohol content typically had better ratings. (Reminder: 
alcohol lowers the density of the wine, which is mostly water.)

Finally, a negative trend seems to exist between `quality` and `chlorides`. 
More chlorides seems to negatively impact quality. I found some information to 
corroborate this observation (see resources text file).

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Yes, there were several.

`pH` and `fixed.acidity` had a modest negative correlation (r = -0.4259). The 
relationship suggests that the higher the `fixed.acidity`, the lower the `pH`. 
That is not an implausable correlation. The more acid in a solution, all else 
equal, the lower the pH of that solution will be.

With that in mind, there is more than one kind of acid in wine that 
contributes to the final pH of a wine. `fixed.acidity`  in this data set 
represents tartaric acid, one of the main acid constituents in wine, 
potentially explaining why we have a noticable trend. But because it is 
not the only acid present, a stronger relationship is not observed.

At first glance it may seem curious that there isn't even a modest 
relationship between `volatile.acidity` (acetic acid content) and `pH`. It may 
also seem strange that no significant relationship appears to exist between 
`citric.acid` and `pH`.

If you look at the measurements of all three acids, however, the above patterns 
actually do make sense. All of the acids are measured in grams per decimeter 
cubed.  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Fixed acidity summary", "\n")
print(summary(white.wine.df$fixed.acidity))

cat("\n", "Volatile acidity summary", "\n")
print(summary(white.wine.df$volatile.acidity))

cat("\n", "Citric acid summary", "\n")
print(summary(white.wine.df$citric.acid))
```

`fixed.acidity` quantities are much higher than either `volatile.acidity` or 
`citric.acid` contents, and they all have comprable pKa ranges (a measure of 
acidity), meaning that `fixed.acidity` (tartaric acid) has greater 
influence on pH.

`total.sulfur.dioxide` and `free.sulfur.dioxide` have a positive correlation 
(r = 0.6155). This makes sense given my previous explanation of their 
relationship (see previous question).

A strong positive relationship exists between `density` and `residual.sugar` 
(r = 0.8390). This makes sense because dissolved solids increase the density 
of aqueous solvents (wine is mostly water, whose density is 1 g / cm^3).

A strong negative relationship exists between `density` and `alcohol` 
(r = -0.7801). This makes sense for the opposite reason. Ethanol has a 
density of 0.7893 g / cm^3, so its presence would decrease the density of 
a wine.

`alcohol` and `residual.sugar` have a moderate negative correlation (r = 
-0.4506). Because alcohol is generated by the fermentation of sugar, if 
fermentation is limited by the winemaker more sugar and less alcohol would 
be in the final wine, all else equal. I believe a stronger relationship is 
not seen because the amount of sugar at the beginning of fermentation is 
different amongst wines. This depends on the grapes' qualities and if sugar is 
added to the starting product.

`density` and `total.sulfur.dioxide` share a moderate positive correlation 
(r = 0.5299). I am unsure why this correlation exists since the density of 
sulfur dioxide is lower than water (0.00293 g/cm^3).

### What was the strongest relationship you found?

The strongest relationship amongst numerical variables was between `density` 
and `residual.sugar` (r = 0.8390).

I do not have a quantification for the strongest relationship between 
`quality` and another feature; however, visually the relationship between
`quality` and `alcohol` seems to be the most dramatic.

***

# Multivariate Plots Section
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = (4 * round(density / 4, digits = 4)), 
                                 y = alcohol)) + 
  geom_line(aes(color = quality.category), stat = "summary", fun.y = median, 
            size = 1) + 
  coord_cartesian(x = c(0.985, 1.010)) +
  labs(title = "Alcohol vs. Density by Quality Category\n(Binned)", 
       x = "Density (Binned)", y = "Alcohol") + 
  theme(panel.background = element_rect(fill="white", color="white"),
        panel.grid.major.y = element_line(color = "gray70"),
        panel.grid.major.x = element_line(color = "gray70"),
        panel.grid.minor.x = element_blank(), 
        plot.title = element_text(size=18, hjust=0.5)) + 
  scale_colour_brewer(palette = "Set1", direction = -1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = (5 * round(pH / 5, digits = 2)), y = volatile.acidity), 
       data = white.wine.df) + 
  geom_line(aes(color = quality), stat = "summary", fun.y = median, size = 1) +
  labs(title = "Volatile acidity vs. pH by Quality Category\n(Binned)", 
       x = "pH (Binned)", y = "Volatile Acidity") + 
  theme(panel.background = element_rect(fill="white", color="white"),
        panel.grid.major.y = element_line(color = "gray70"),
        panel.grid.major.x = element_line(color = "gray70"),
        panel.grid.minor.x = element_blank(), 
        plot.title = element_text(size=18, hjust=0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = (2 * round(pH / 2, digits = 2)), y = volatile.acidity), 
       data = white.wine.df) + 
  geom_point(aes(color = quality)) +
  labs(title = "Volatile acidity vs. pH by Quality Category\n(Binned)", 
       x = "pH (Binned)", y = "Volatile Acidity") + 
  theme(panel.background = element_rect(fill="white", color="white"),
        panel.grid.major.y = element_line(color = "gray70"),
        panel.grid.major.x = element_line(color = "gray70"),
        panel.grid.minor.x = element_blank(), 
        plot.title = element_text(size=18, hjust=0.5))
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

The first multivariate plot is of `alcohol`,`density`, and `quality.category`. 
As seen before, `alcohol` and `density` are strongly related to one another: 
regardless of a wine's category, `alcohol` and `density` are negatively 
correlated. Also from earlier exploration, the relationship between 
`quality.category` and `alcohol` still holds true. The "Best" category wines 
tend to have higher `alcohol` than their peers at the same `density` when 
`density` is less than 0.9955 g/cm^3.

I also tried plotting `volatile.acidity`, `pH` and `quality` two different 
ways: a line plot and a scatterplot form. Each brought different insights.

The line plot showed that the worse quality wine (ratings 3 and 4) tended 
toward higher `volatile.acidity`, regardless of pH (though there were some 
spikes). The middle rated wines (5, 6, 7) can be seen closer to the bottom 
(lower `volatile.acidity`) and also spanning the largest pH range. The highest 
rating category in this data set (9) spans the shortest pH range.

The scatterplot makes it unmistakably clear that the middle rated wines are 
the most abundant and span across the widest pH scale. It is also possible to 
see that wines with higher `volatile.acidity` are a mix of the worst rated 
wines and middle rated wines; however, worse rated wines tend to appear with 
higher rather than lower `volatile.acidity`.

### Were there any interesting or surprising interactions between features?

I did not see anything new that was not already seen in the bivariate 
analysis.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

N/A

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
fa.iqr <- quantile(white.wine.df$fixed.acidity, 0.75) - quantile(white.wine.df$fixed.acidity , 0.25)
fa.no.outliers <- subset(white.wine.df, white.wine.df$fixed.acidity < 
                       (quantile(white.wine.df$fixed.acidity, 0.75) + 3*fa.iqr))

ggplot(data = fa.no.outliers, aes(x = fixed.acidity, y = pH)) +
  geom_point(color = "dodgerblue4", alpha = 0.2, position = "jitter") +
  geom_smooth(method = "lm", color = "navy") +
  labs(x = "Fixed acidity (g / dm^3)", y = "pH", title = "pH by Fixed Acidity") + 
  scatterplot.theme
```

### Description One
This scatterplot depicts the relationship between `pH` and `fixed.acidity` 
(r = -0.4259). Although it is only a moderate relationship, the 
inquiry that followed was interesting because I ended up learning about the 
acid composition of wine.

While it is not surprising that greater 
`fixed.acidity` correlates to lower pH, the imperfect relationship suggested 
that something more was at play. As discussed in the __Bivarate Analysis__ 
section, `fixed.acidity` only represents one of the several acids present in 
wine (tartaric acid). After some brief investigation, it was clear that other 
acids, like malic acid (not measured in this data set), also contribute to 
the final pH of a wine. When I checked `volatile.acidity` (acetic acid) and 
`citric.acid` for correlation to pH, the relationships were not strong, meaning 
that there was much less of either acid present compared to tartaric acid. 

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(data = white.wine.df, aes(x = quality, y = alcohol)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "Alcohol\n(% by Volume)", 
       title = "Alcohol Content by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme + 
  guides(fill = guide_legend(reverse=TRUE))
```

### Description Two
This boxplot shows alcohol content of wines by quality category. I elected to 
group the wines into "Worst", "Middle", and "Best" for visualization purposes.

This plot was particularly interesting not simply because it showed a clear 
preference for higher alcohol wines, but because it revealed a market trend. 
As discussed in the __Bivarate Analysis__ section, New World wines with 
bolder, "riper" flavors are in vogue, which consequently come with higher 
alcohol content.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(aes(x = (5 * round(pH / 5, digits = 2)), y = volatile.acidity), 
       data = white.wine.df) + 
  geom_line(aes(color = quality), stat = "summary", fun.y = median, size = 1) +
  labs(title = "Volatile acidity vs. pH by Quality Category\n(Binned)", 
       x = "pH (Binned)", y = "Volatile Acidity") + 
  theme(panel.background = element_rect(fill="white", color="white"),
        panel.grid.major.y = element_line(color = "gray70"),
        panel.grid.major.x = element_line(color = "gray70"),
        panel.grid.minor.x = element_blank(), 
        plot.title = element_text(size=18, hjust=0.5))
```

### Description Three
Although this line plot may seem somewhat noisy, it actually contains a lot 
of information of use. The finer details are discussed in the 
__Multivariate Analysis__ section, but overall it shows a slight trend in 
quality and volatile acidity. It also shows that no significant relationship 
exists between pH and volatile acidity (which reinforces previous findings).

At a bigger picture level, this plot demonstrates that more is at work when 
it comes to quality rating: there are peaks and valleys for the different 
quality ratings, meaning some wines certainly buck the trend for 
`volatile.acidity`. It also shows the diversity in pH for wines. These 
observations reinforce my general impression that quality is influenced by 
multiple chemical properties of wine; that no one property is the ultimate 
dictator of quality.

------

# Reflection

This was a rewarding data set to work on from both a professional and personal 
perspective. Going through a full analysis of the data, starting at a 
univariate level, progressing through a bivariate level, then finishing 
at a multivariate level, gave me the opportunity to apply and polish what 
I have learned. What I especially enjoyed was customizing plots. I was 
able to learn much more about ggplot2 in the process: the different layers 
available, some of their respective intricacies, and the various ways I could 
portray my findings.

From a personal perspective, I loved learning more about wine! My academic 
background is in chemistry and I have a particular fondness for natural 
product chemistry and food chemistry. 

Although I wouldn't deem them "struggles" exactly, having to rummage the 
internet and other resources (I have a copy of "R in a Nutshell") for code
solutions took time and effort. For example, getting a good grasp of the 
`theme()` layer took a good amount of investigation. 

Another challenge was making sense of some of the observations resulting from 
my queries. This was the fun part though! I never knew that "New World" wines 
were a class of wines, nor that they were defined by having riper, fruitier 
flavors compared to "Old World" wines. It was also interesting to learn that 
these in fashion wines are fermented with riper grapes (more sugar rich) 
leading to more alcoholic wines -- explaining my original observation that 
wines with higher `quality` ratings tended to have higher `alcohol`.

I was also suprised by my findings for `quality` and `chlorides`. I had no 
clue that wines could be "salty" to a detectable level, though it makes 
sense that these wines would have a lower `quality`. I don't think I would 
seek out or enjoy a noticeably salty wine!

In the future, once I have learned more about predictive modeling, I think 
building a model could be very interesting. It would be a great way to 
test my present findings and also to get a better understanding of what people 
look for in wines. Another thing that would interesting would be repeating 
the experiment and adding more variables -- e.g. all the major acids like 
malic, lactic, succinic acid, etc, as well as an oxidation marker of some 
sort and acetaldehyde levels. Many of these have profound effects on 
the sensory experience of wine. So we may be missing important information 
if these are not included.