WHITE WINE QUALITY ANALYSIS by Kasey Cox
========================================================

__Overview of data to be analyzed:__ This tidy data set contains 4,898 white 
wines with 11 variables on quantifying the chemical properties of each wine. At 
least 3 wine experts rated the quality of each wine, providing a rating between 
0 (very bad) and 10 (very excellent). This is a curated data set provided by
Udacity using the following research article:

Cortez, P.; Cerdeira, A.; Almeida, F.; Matos, T.; Reis, J.  
Modeling wine preferences by data mining from physicochemical properties.  
_Decision Support Systems_. __2009__, _47_, 547-553.

[Elsevier](http://dx.doi.org/10.1016/j.dss.2009.05.016)  
[Pre-press (pdf)](http://www3.dsi.uminho.pt/pcortez/winequality09.pdf)  
[bib](http://www3.dsi.uminho.pt/pcortez/dss09.bib)

__Question guiding investigation:__ Which chemical properties influence the 
quality of white wines?

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

library(ggplot2)
library(gridExtra)
library(plyr)
library(GGally)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load the Data
white.wine.df <- read.csv("wineQualityWhites.csv")
```

***

# Univariate Plots Section

### Preliminary information on the data set  
```{r echo=FALSE, message=FALSE, warning=FALSE}
str(white.wine.df)
summary(white.wine.df)
```

### Organic acid contents and overall acidity (pH)  
```{r echo=FALSE, message=FALSE, warning=FALSE}
ta.hist <- ggplot(data = white.wine.df, aes(x = fixed.acidity)) +
  geom_histogram() +
  labs(x = "Fixed acidity (g/dm^3 of tartaric acid)")

ca.hist <- ggplot(data = white.wine.df, aes(x = citric.acid)) +
  geom_histogram() +
  labs(x = "Citric acid content (g/dm^3)")

aa.hist <- ggplot(data = white.wine.df, aes(x = volatile.acidity)) +
  geom_histogram() +
  labs(x = "Volatile Acidity (g/dm^3 of acetic acid)")

ph.hist <- ggplot(data = white.wine.df, aes(x = pH)) +
  geom_histogram() +
  labs(x = "pH")

grid.arrange(ta.hist, ca.hist, aa.hist, ph.hist, ncol =2)
```

The various acid measures all have a normal distribution, with all but pH having
a few outliers (in a visual sense).

__Citric acid:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}

print(paste("A few of the more extreme values...", head(subset(white.wine.df, 
            white.wine.df$citric.acid > 0.9, select = citric.acid))))

print(paste("3rd quartile:", quantile(white.wine.df$citric.acid, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$citric.acid, 0.99)))

```

The mean (0.3342) and median (0.3200) are similar in the citric acid 
distribution, meaning few wines exist in the tail and that most of the wines 
follow the normal distribution.

__Fixed acidity (tartaric acid):__  
```{r echo=FALSE, message=FALSE, warning=FALSE}
#subset(white.wine.df, 
       #white.wine.df$fixed.acidity > 10.0)[c("X", "fixed.acidity")]

print(paste("A few of the more extreme values...", head(subset(white.wine.df, 
            white.wine.df$fixed.acidity > 10.0, select = fixed.acidity))))

print(paste("3rd quartile:", quantile(white.wine.df$fixed.acidity, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$fixed.acidity, 0.99)))
```

The mean (6.855) and median (6.800) of fixed acidities are close in the 
distribution, meaning few wines exist in the tail and that most of the wines 
follow the normal distribution.

__Volatile acidity (acetic acid):__  
```{r echo=FALSE, message=FALSE, warning=FALSE}

print(paste("A few of the more extreme values...", head(subset(white.wine.df, 
            white.wine.df$volatile.acidity > 0.9, select = volatile.acidity))))

print(paste("3rd quartile:", quantile(white.wine.df$volatile.acidity, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$volatile.acidity, 0.99)))
```

The mean (0.2782) and median (0.2600) of volatile acidities are close in the 
distribution, meaning few wines exist in the tail and that most of the wines 
follow the normal distribution.

### Sulfur measures
```{r Sulfur measures, echo=FALSE, message=FALSE, warning=FALSE}
sulf.hist <- ggplot(data = white.wine.df, aes(x = sulphates)) +
  geom_histogram() +
  labs(x = "Sulphates (added potassium sulphate - g/dm^3)")

free.sulf.diox <- ggplot(data = white.wine.df, aes(x = free.sulfur.dioxide)) +
  geom_histogram() +
  labs(x = "free sulfur dioxide (mg / dm^3)")

total.sulf.diox <- ggplot(data = white.wine.df, aes(x = total.sulfur.dioxide)) +
  geom_histogram() +
  labs(x = "total sulfur dioxide (mg / dm^3)")

grid.arrange(sulf.hist, free.sulf.diox, total.sulf.diox, ncol = 2)
```

The sulfur measures also have normal distribution, though an extreme outlier
exists for the free sulfur dioxide histogram.

__Sulfates (added):__  
The distribution of sulfates is roughly normal.  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
summary(white.wine.df$sulphates)

print(paste("3rd quartile:", quantile(white.wine.df$sulphates, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$sulphates, 0.99)))
```

__Free sulfur dioxide:__  
The distribution of free sulfur dioxide is roughly normal, with an extreme
outlier (max).
```{r, echo=FALSE, message=FALSE, warning=FALSE}
summary(white.wine.df$free.sulfur.dioxide)

print(paste("3rd quartile:", quantile(white.wine.df$free.sulfur.dioxide, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$free.sulfur.dioxide, 0.99)))
```

__Total sulfur dioxide:__  
The distribution of total sulfur dioxide is roughly normal, with an extreme
outlier (max).
```{r, echo=FALSE, message=FALSE, warning=FALSE}
summary(white.wine.df$total.sulfur.dioxide)

print(paste("3rd quartile:", quantile(white.wine.df$total.sulfur.dioxide, 0.75)))

print(paste("99th percentile:", 
            quantile(white.wine.df$total.sulfur.dioxide, 0.99)))
```


### Other wine attributes
__Density and residual sugar:__  
```{r echo=FALSE, message=FALSE, warning=FALSE}
res.sugar.hist <- ggplot(data = white.wine.df, aes(x = residual.sugar)) +
  geom_histogram() +
  labs(x = "Residual sugars (g/dm^3)")

density.hist <- ggplot(data = white.wine.df, aes(x = density)) +
  geom_histogram() +
  labs(x = "density (g/cm^3)")

grid.arrange(res.sugar.hist, density.hist, ncol = 2)
```

In the distributions for both density and residual sugar, we can see an extreme 
outlier (maybe more than one) in both. The distributions are otherwise 
relatively narrow. My suspicion is that they are the same wine (or few wines). 
Dissolved sugar increases the density of water (which is 1.00 g/cm^3).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(paste("High density wine:", 
            white.wine.df$density[white.wine.df$density > 1.02]))

print(paste("High residual sugar wine:", 
            white.wine.df$residual.sugar[white.wine.df$residual.sugar > 60]))

subset(white.wine.df, white.wine.df$density > 1.02)[c("residual.sugar", 
                                                      "density")]
```

We can see that they are indeed the same wine.

__Sodium chloride:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
nacl.hist <- ggplot(data = white.wine.df, aes(x = chlorides)) +
  geom_histogram() +
  labs(x = "Sodium chloride (g/dm^3)")

summary(white.wine.df$chlorides)
```

There is a little bit of a tail in the sodium chloride data; most of the 
values are less than 0.05, yet there are a handful of values greater than 0.2 
Let's do a log base 10 transform on the x-axis.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
nacl.log10.hist <- ggplot(data = white.wine.df, aes(x = chlorides)) +
  geom_histogram() +
  labs(x = "Log 10 Sodium Chloride") +
  scale_x_log10()

grid.arrange(nacl.hist, nacl.log10.hist, ncol = 1)
```

We see that there really aren't many data points beyond 0.1 g/dm^3. This is
consistent with the mean not being far from the median.

__Alcohol (ABV - % by volume):__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = alcohol)) +
  geom_histogram(binwidth = 0.25) +
  labs(x = "Alcohol (ABV - % by volume)")
```

The distribution of ABV is roughly normal.

__Quality Ratings:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Quality Rating (0 - 10)") +
  scale_x_continuous(limits = c(2, 10), breaks = seq(0, 10, 1))
```

The quality ratings are uniformly distributed around 6 (the mode).

### More about the best, middle, and worst wines
Wine ratings were made on a scale of 0 to 10, with 0 being the worst. In this
data, wines ranged from 3 to 9. I divided the ratings as follows:  
* Worst: wines rated 3, 4  
* Middle: wines rated 5, 6, 7  
* Best: wines rated 8, 9  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Make quality categorical
white.wine.df$quality <- factor(x = white.wine.df$quality, levels = c(0:10), ordered = TRUE)

# Add new categorical variable
white.wine.df$quality.category <- ifelse(white.wine.df$quality >= 8, "Best", 
                                         ifelse(white.wine.df$quality >= 5, 
                                                "Middle", 
                                                ifelse(white.wine.df$quality 
                                                       >= 3, "Worst", "Other")))

# Number of wines in each category (need plyr loaded)
print(paste("Best rated wines:", 
            count(white.wine.df$quality.category == "Best")))
print(paste("Middle rated wines:", 
            count(white.wine.df$quality.category == "Middle")))
print(paste("Worst rated wines:", 
            count(white.wine.df$quality.category == "Worst")))
```

I am going to focus on some of the features of the best wines.

```{r}
best.va.hist <- ggplot(data = subset(white.wine.df, quality.category == "Best"), 
                       aes(x = volatile.acidity)) + 
  geom_histogram(bins = 40)

best.ph.hist <- ggplot(data = subset(white.wine.df, quality.category == "Best"), 
                       aes(x = pH)) +
  geom_histogram(bins = 40)

best.fsd.hist <- ggplot(data = subset(white.wine.df, quality.category == "Best"), 
                        aes(x = free.sulfur.dioxide)) +
  geom_histogram(bins = 40)

best.rs.hist <- ggplot(data = subset(white.wine.df, quality.category == "Best"), 
                       aes(x = residual.sugar)) +
  geom_histogram(bins = 40)

best.alcohol.hist <- ggplot(data = subset(white.wine.df, 
                                          quality.category == "Best"), 
                            aes(x = alcohol)) +
  geom_histogram(bins = 40)

best.nacl.hist <- ggplot(data = subset(white.wine.df, quality.category == "Best"), 
                         aes(x = chlorides)) +
  geom_histogram(bins = 40)

grid.arrange(best.va.hist, best.ph.hist, best.fsd.hist, best.rs.hist, best.alcohol.hist, best.nacl.hist, ncol = 2)
```

It appears that in the best wines (rated 8 or greater), certain features show
less variance whereas others vary greatly. For example, chloride content does 
not seem to vary much (except for an extreme outlier); residual sugar has a 
lot of variation.

Most of the distribution show a normal or somewhat normal distribution. The
chloride distribution is relatively normal. The residual sugar distribution
seems almost uniform in its distribution.

# Univariate Analysis

### What is the structure of your dataset?
This data has 4898 observations (wines) with 13 features (variables). These 
features are "X", "fixed.acidity", "volatile.acidity", "citric.acid", 
"residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", 
"density", "pH", "sulphates", "alcohol", and "quality".

Most of the variables are numerical, except for "quality". Although it consists
of numbers, the numbers represent rank rather than being numbers in the strict 
mathematical sense. When I loaded the data, R interpreted it as a numerical 
vector, so I purposely changed it to a factor to properly reflect what it 
stands for.

### What is/are the main feature(s) of interest in your dataset?

The main feature of my data set is "quality" which, as mentioned in the previous
answer, represets a ranking: 0 to 10 with 0 being the worst quality rating and 
10 being the best quality rating.

The goal of this analysis is to determine which chemical properties affect the
wine's rating (quality).

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Because we want to know which chemical properties affect quality, a feature's 
importance will be determined relative to its effect on quality.

So far, a few features stand out: alchohol content, chloride content, free 
sulfur dioxide, and volatile acidity.

### Did you create any new variables from existing variables in the dataset?

Yes. I created a "quality.category" (factor) variable using the "quality" 
variable. This factor has three levels: "Worst", "Middle", and "Best". If a 
wine was rated poorly (0-3), its quality.category value is "Worst". A rating 
of 5-7 corresponds to "Middle". A rating of 8-10 represents "Best".

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

The histogram of chloride content had a long tail with what appeared to be many 
data points along its length. I did a log transformation along its x-axis to 
see if there really were a good amount of data points in the tail. After the 
transformation, I found that the tail was less significant than I had 
originally thought.

***

# Bivariate Plots Section

### Preliminary plotting
For my preliminary bivariate plotting, I am using the `ggpairs` function to 
help guide my later, targeted plots. The resulting plot matrix will give 
me an initial look at any possible relationships that may be worth exploring 
further.

To prevent overcrowded plot matrices, I will be subsetting the data into 
different groups based on my hunches about variables that may have something 
to do with one another. I will also create a subset that includes variables
that are, at least to my understanding, seemingly unrelated.

__Acidity subset__  
```{r, message=FALSE, warning=FALSE}
acid.subsetted.data <- subset(white.wine.df, !is.na(white.wine.df), 
                                 select = c(fixed.acidity, volatile.acidity, 
                                            citric.acid, pH, quality))

ggpairs(acid.subsetted.data, 
        lower = list(continuous = wrap("points")))
```

There only appears to be one moderate correlation amongst the numeric variables 
in this subset: between `pH` and `fixed.acidity` (r = -0.426). 

There also only appears to be one correlation between `quality` (categorical 
variable) and `pH`. This will require further exploration below.

__Sulfur subset__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
sulfur.subsetted.data <- subset(white.wine.df, !is.na(white.wine.df), 
                                 select = c(free.sulfur.dioxide, 
                                            total.sulfur.dioxide, 
                                            sulphates, quality))

ggpairs(sulfur.subsetted.data, 
        lower = list(continuous = wrap("points")))
```

Amongst the numerical variables, a correlation exists between 
`free.sulfur.dioxide` and `total.sulfur.dioxide` (r = 0.616). 

For `quality`, the only relationship that appears to exist is between `quality` 
and `total.sulfur.dioxide`, but this requires more investigation below.

__Other features subset__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
other.subsetted.data <- subset(white.wine.df, !is.na(white.wine.df), 
                                 select = c(residual.sugar, chlorides, 
                                            density, alcohol, quality))

ggpairs(other.subsetted.data, 
        lower = list(continuous = wrap("points")))
```

A strong positive correlation exists between `residual.sugar` and `density` 
(r = 0.839). There is also a strong negative correlation between `alcohol` 
and `density` (r = -0.78).  Finally, there is a modest negative correlation 
between `alcohol` and `residual.sugar` (r = -0.4506).

As for relationships involving quality, there definitely appears to be a 
positive relationship between `quality` and `alcohol`. There also appears to 
be negative relationships between `quality` and `chlorides` as well as 
`quality` and `density`. A relationship between `quality` and `residual.sugar` 
seems unclear. All of these will be explored further below.

__Random features subset__
```{r, echo=FALSE, message=FALSE, warning=FALSE}
random.subsetted.data <- subset(white.wine.df, !is.na(white.wine.df), 
                                 select = c(sulphates, pH, 
                                            total.sulfur.dioxide, 
                                            density, citric.acid, quality))

ggpairs(random.subsetted.data, 
        lower = list(continuous = wrap("points")))
```

A modest positive correlation seems to exists between `density` and 
`total.sulfur.dioxide` (r = 0.53).

No other significant correlations seems to exist amongst the numerical 
variables.

No new relationships involving `quality` are obvious.

### Relationships amongst chemical properties
Taking a close look at above findings between different pairs of chemical 
properties (based on correlation coefficient findings).

```{r, message=FALSE, warning=FALSE}
scatterplot.theme <- theme(panel.background = element_rect(fill="white", 
                                                       color="white"),
        panel.grid.major.y = element_line(color = "gray70"),
        panel.grid.major.x = element_line(color = "gray70"),
        panel.grid.minor.x = element_blank(), 
        plot.title = element_text(size=18, hjust=0.5),
        legend.title = element_blank())
```


__pH and Fixed Acidity__  
[With 4 "far out" outliers eliminated](https://en.wikipedia.org/wiki/Outlier#Tukey.27s_test)
```{r, echo=FALSE, message=FALSE, warning=FALSE}
fa.iqr <- quantile(white.wine.df$fixed.acidity, 0.75) - quantile(white.wine.df$fixed.acidity , 0.25)
fa.no.outliers <- subset(white.wine.df, white.wine.df$fixed.acidity < 
                       (quantile(white.wine.df$fixed.acidity, 0.75) + 3*fa.iqr))

ggplot(data = fa.no.outliers, aes(x = fixed.acidity, y = pH)) +
  geom_point(color = "dodgerblue4", alpha = 0.2, position = "jitter") +
  geom_smooth(method = "lm", color = "navy") +
  labs(x = "Fixed acidity (g / dm^3)", y = "pH", title = "pH by Fixed Acidity") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(paste("r =", cor(x = white.wine.df$fixed.acidity, 
                       y = white.wine.df$pH, method = "pearson")))
```

__Total sulfur dioxide and free sulfur dioxide__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide)) +
  geom_point(color = "dodgerblue4", alpha = 0.1, position = "jitter") +
  geom_smooth(method = "lm", color = "navy") +
  coord_cartesian(xlim = c(0,150)) +
  labs(x = "Free Sulfur Dioxide\n(mg / dm^3)", y = "Total Sulfur Dioxide\n(mg / dm^3)", title = "Total vs. Free Sulfur Dioxide") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(paste("r =", cor(x = white.wine.df$free.sulfur.dioxide, 
                       y = white.wine.df$total.sulfur.dioxide, 
                       method = "pearson")))
```

__Density vs. residual sugar__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = residual.sugar, y = density)) +
  geom_point(color = "dodgerblue4", alpha = 0.1, position = "jitter") +
  geom_smooth(method = "lm", color = "navy") +
  coord_cartesian(xlim = c(0, 35), ylim = c(0.985, 1.015)) +
  labs(x = "Residual Sugar\n(g / dm^3)", y = "Density\n(g / cm^3)", 
       title = "Density vs. Residual Sugar") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(paste("r =", cor(x = white.wine.df$residual.sugar, 
                       y = white.wine.df$density, 
                       method = "pearson")))
```

__Alcohol vs. residual sugar__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = residual.sugar, y = alcohol)) +
  geom_point(color = "dodgerblue4", alpha = 0.2, position = "jitter") +
  geom_smooth(method = "lm", color = "navy", se = FALSE, formula = y ~ poly(x, 2)) +
  coord_cartesian(x = c(0, 35)) +
  labs(x = "Residual Sugar\n(g / dm^3)", y = "Alcohol\n(% by Volume)", 
       title = "Alcohol vs. Residual Sugar") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(paste("r =", cor(x = white.wine.df$residual.sugar, 
                       y = white.wine.df$alcohol, 
                       method = "pearson")))
```

__Density vs. alcohol__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = alcohol, y = density)) +
  geom_point(color = "dodgerblue4", alpha = 0.1, position = "jitter") +
  geom_smooth(method = "lm", color = "navy") + 
  coord_cartesian(ylim = c(0.985, 1.015)) + 
  labs(x = "Alcohol\n(% by volume)", y = "Density\n(g / cm^3)", 
       title = "Density vs. Alcohol") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(paste("r =", cor(x = white.wine.df$alcohol, 
                       y = white.wine.df$density, 
                       method = "pearson")))
```

__Density and Total Sulfur Dioxide__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = total.sulfur.dioxide, y = density)) +
  geom_point(color = "dodgerblue4", alpha = 0.1, position = "jitter") +
  geom_smooth(method = "lm", color = "navy") + 
  coord_cartesian(ylim = c(0.985, 1.015)) + 
  labs(x = "Total Sulfur Dioxide\n(mg / dm^3)", y = "Density\n(g / cm^3)", 
       title = "Density vs. Total Sulfur Dioxide") + 
  scatterplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(paste("r =", cor(x = white.wine.df$total.sulfur.dioxide, 
                       y = white.wine.df$density, 
                       method = "pearson")))
```


### Relationships between quality and other chemical properties
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Boxplot theme
boxplot.theme <- theme(panel.background = element_rect(fill="gray90", 
                                                       color="gray90"),
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(color = "gray50"),
        panel.grid.minor.x = element_blank(), 
        plot.title = element_text(size=18, hjust=0.5),
        legend.title = element_blank())

boxplot.colors <- scale_fill_manual(values=c("green3", "deepskyblue3", "red2"))
```

__Volatile acidity:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, y = volatile.acidity)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "Volatile Acidity\n(g/dm^3)",
       title = "Volatile Acidity by Quality") +
  boxplot.colors +
  boxplot.theme +
  stat_summary(fun.y = mean, geom = "point", shape = 4)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median volatile acidity for 'Best' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == "Best")$volatile.acidity)

cat("\nMedian volatile acidity for 'Middle' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == "Middle")$volatile.acidity)

cat("\nMedian volatile acidity for 'Worst' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == "Worst")$volatile.acidity)
```

It seems that poorly rated wines tend to have higher volatile acidity.

__pH:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, y = pH)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "pH", title = "pH by Quality") +
  boxplot.colors +
  boxplot.theme +
  stat_summary(fun.y = mean, geom = "point", shape = 4)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median pH for 'Best' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == "Best")$pH)

cat("\nMedian pH for 'Middle' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == "Middle")$pH)

cat("\nMedian pH for 'Worst' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == "Worst")$pH)
```

Better wines appear to have a (relatively) higher pH, with the middle rated
wines having a lot of variability.

__Free sulfur dioxide:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, 
                                 y = free.sulfur.dioxide)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip(ylim = c(0, 150)) +
  labs(x = "Quality Category", y = "Free Sulfur Dioxide\n(mg / dm^3)", 
       title = "Free Sulfur Dioxide by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median Free Sulfur Dioxide for 'Best' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Best")$free.sulfur.dioxide)

cat("\nMedian Free Sulfur Dioxide for 'Middle' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Middle")$free.sulfur.dioxide)

cat("\nMedian Free Sulfur Dioxide for 'Worst' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Worst")$free.sulfur.dioxide)

cat("\nMax Free Sulfur Dioxide\n")
subset(white.wine.df, 
       white.wine.df$free.sulfur.dioxide > 150)[c("free.sulfur.dioxide", 
                                                  "quality", 
                                                  "quality.category")]
```

The median free sulfur dioxide is lower in the poorly rated wines; however, 
there is a lot variability amongst these wines. Note that the wine with the 
highest free sulfur dioxide is in the worst category.

__Alcohol (ABV - % by volume):__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, y = alcohol)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip() +
  labs(x = "Quality Category", y = "Alcohol\n(ABV)", 
       title = "ABV by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median Alcohol for 'Best' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Best")$alcohol)

cat("\nMedian Alcohol for 'Middle' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Middle")$alcohol)

cat("\nMedian alcohol for 'Worst' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Worst")$alcohol)
```

The better rated wines generally have higher alcohol content.

__Density:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, y = density)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip(ylim = c(0.985, 1.015)) +
  labs(x = "Quality Category", y = "Density\n(g/cm^3)", 
       title = "Density by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median Density for 'Best' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Best")$density)

cat("\nMedian Density for 'Middle' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Middle")$density)

cat("\nMedian Density for 'Worst' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Worst")$density)
```

Better rated wines tend to have a lower density.

__Chlorides:__  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality.category, y = chlorides)) +
  geom_boxplot(aes(fill = quality.category)) +
  coord_flip(ylim = c(0, 0.15)) +
  labs(x = "Quality Category", y = "Sodium chloride\n(g / dm^3)", 
       title = "Sodium Chloride Content by Quality") +
  stat_summary(fun.y = mean, geom = "point", shape = 4) +
  boxplot.colors +
  boxplot.theme
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Median Chlorides for 'Best' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Best")$chlorides)

cat("\nMedian Chlorides for 'Middle' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Middle")$chlorides)

cat("\nMedian Chlorides for 'Worst' wines\n")
median(subset(white.wine.df, white.wine.df$quality.category == 
                "Worst")$chlorides)
```

The worse rated wines tend to have higher sodium chloride content.

***

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

I observed several relationships involving the main feature (`quality`). Poorly 
rated wines ("Worst" `quality.category`) tended to have higher 
`volatile.acidity`. In the text accompanying the data set, there is 
information on the variables contained in the data. Of note is this: "volatile 
acidity: the amount of acetic acid in wine, which at too high of levels can 
lead to an unpleasant, vinegar taste." This would explain why at higher values 
these wines would end up in the worst category. Perhaps "Middle" and "Best" 
wines do not hit the threshold of "too high of levels", thereby not 
experiencing any negative impact from `volatile.acidity`.

Wines with a higher `pH` value tended to have a better `quality` rating. This 
holds true for all three `quality.category` subsets ("Worst", "Middle", 
"Best"). I do _not_ believe this is because wines with higher 
`volatile.acidity` generally have worse ratings. In the __Acidity subset__ 
plot matrix above the __pH__ boxplot I go into a discussion about how 
`volatile.acidity` has a low impact on final `pH` (how, in fact, it is 
`fixed.acidity` that largely determines final `pH`). Maybe it is a tasting 
preference to have slightly less acidic wine (they are all acidic 
nonetheless).

A relationship also appears to exist between `quality` and 
`free.sulfur.dioxide`. Another note from the text accompanying the data set 
states: "total sulfur dioxide: amount of free and bound forms of S02; in low 
concentrations, SO[2] is mostly undetectable in wine, but at free SO[2] 
concentrations over 50 ppm, SO[2] becomes evident in the nose and taste 
of wine"

Somewhat counterintuitively, The median free sulfur dioxide is lower in the 
poorly rated wines; however, there is a lot variability amongst these wines. 
Note that the wine with the highest free sulfur dioxide is in the worst 
category. A closer examination reveals that at the cited 50 ppm, that 
would mean free sulfur dioxide would have to be at least 49.9 mg/dm^3 to be 
detectable. Most wines (in all categories) are less than this threshold. So 
for the most part, too much free SO2 likely isn't a factor unless it is at 
unreasonably high amounts. So what else could explain this trend?

I did a little searching to make sense of this observation ([Sulfur dioxide management](http://extension.psu.edu/food/enology/wine-production/wine-made-easy-fact-sheets/sulfur-dioxide-management)). 
Sulfur dioxide is 
an important _preservative_ and _stabilizer_ used world-wide in wine making. It 
makes sense then that worse quality wines may not be well preserved -- 
developing off flavors, etc. Obviously, SO2 is not the only factor affecting 
quality. A poorly made wine, no matter how well preserved, will still get bad 
ratings.

`quality` ratings tend to be higher when `alcohol` is higher. This appears to 
be a preference. According to an article in the 
[Scientific American](https://www.scientificamerican.com/article/wine-becomes-more-like-whisky-as-alcohol-content-gets-high/:)  
> The recent surge in wine's punch [alcohol content] is largely a result, 
scientists say, of a fashion for deeply colored wines with fewer “green” 
qualities and more bright, ripe, fruity flavors. As New World wines in this 
style have drawn more fans, even European winemakers accustomed to making 
lower-alcohol wines in less ripe styles are beginning to follow suit. But 
producing wines with those flavors means letting grapes hang longer on the 
vine, and with longer hang times comes bigger sugar. The more sugar the wine 
yeast _S. cerevisiae_ has to work with, the more alcohol it will make.

Along similiar lines `quality` tends to be higher in wines with lower 
`density`. This is consistent with the previous relationship that showed 
wines with higher alcohol content typically had better ratings. (Reminder: 
alcohol lowers the density of the wine, which is mostly water.)

Finally, a negative trend seems to exist between `quality` and `chlorides`. 
More chlorides seems to negatively impact quality. I found some information to 
corroborate this observation:  
[Peer reviewed paper](http://www.scielo.br/scielo.php?script=sci_arttext&pid=S0101-20612015000100095)  
[Blog post](http://wineoscope.com/2015/10/02/when-a-wine-is-salty-and-why-it-shouldnt-be/)

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Yes, there were several.

`pH` and `fixed.acidity` had a modest negative correlation (r = -0.4259). The 
relationship suggests that the higher the `fixed.acidity`, the lower the `pH`. 
That is not an implausable correlation. The more acid in a solution, all else 
equal, the lower the pH of that solution will be.

With that in mind, there is more than one kind of acid in wine that 
contributes to the final pH of a wine. `fixed.acidity`  in this data set 
represents tartaric acid, one of the main 
[acid constituents in wine](http://waterhouse.ucdavis.edu/whats-in-wine/fixed-acidity), 
potentially explaining why we have a noticable trend. But because it is not 
the only acid present, a stronger relationship is not observed.

At first glance it may seem curious that there isn't even a modest 
relationship between `volatile.acidity` (acetic acid content) and `pH`. It may 
also seem strange that no significant relationship appears to exist between 
`citric.acid` and `pH`.

If you look at the measurements of all three acids, however, the above patterns 
actually do make sense. All of the acids are measured in grams per decimeter 
cubed.  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat("Fixed acidity summary", "\n")
print(summary(white.wine.df$fixed.acidity))

cat("\n", "Volatile acidity summary", "\n")
print(summary(white.wine.df$volatile.acidity))

cat("\n", "Citric acid summary", "\n")
print(summary(white.wine.df$citric.acid))
```

Fixed acidity quantities are much higher than either volatile acidity or citric 
acid contents, and they all have comprable pK[a] ranges (a measure of acidity), 
meaning that fixed acidity (tartaric acid) has greater influence on pH.

`total.sulfur.dioxide` and `free.sulfur.dioxide` have a positive correlation 
(r = 0.6155). This could suggest that `free.sulfur.dioxide` is the majority 
contributor to `total.sulfur.dioxide` (in this data set, 
`total.sulfur.dioxide` is defined as the "amount of free and bound forms of 
S02").

A strong positive relationship exists between `density` and `residual.sugar` 
(r = 0.8390). This makes sense because dissolved solids increase the density 
of aqueous solvents (wine is mostly water, whose density is 1 g / cm^3).

A strong negative relationship exists between `density` and `alcohol` 
(r = -0.7801). This makes sense for the opposite reason. Ethanol has a 
density of 0.7893 g / cm^3, so its presence would decrease the density of 
a wine.

`alcohol` and `residual.sugar` have a moderate negative correlation (r = 
-0.4506). Because alcohol is generated by the fermentation of sugar, if 
fermentation is limited by the winemaker more sugar and less alcohol would 
be in the final wine, all else equal. I believe a stronger relationship is 
not seen because the amount of sugar at the beginning of fermentation is 
different amongst wines. This depends on the grapes' qualities and if sugar is 
added to the starting product.

`density` and `total.sulfur.dioxide` share a moderate positive correlation 
(r = 0.5299). I am unsure why this correlation exists since the density of 
sulfur dioxide is lower than water (0.00293 g/cm^3).

### What was the strongest relationship you found?

The strongest relationship amongst numerical variables was between `density` 
and `residual.sugar` (r = 0.8390).

I do not have a quantification for the strongest relationship between 
`quality` and another feature; however, visually the relationship between
`quality` and `alcohol` seems to be the most dramatic.

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!