WHITE WINE QUALITY ANALYSIS by Kasey Cox
========================================================

__Overview of data to be analyzed:__ This tidy data set contains 4,898 white 
wines with 11 variables on quantifying the chemical properties of each wine. At 
least 3 wine experts rated the quality of each wine, providing a rating between 
0 (very bad) and 10 (very excellent). This is a curated data set provided by
Udacity using the following research article:

Cortez, P.; Cerdeira, A.; Almeida, F.; Matos, T.; Reis, J.  
Modeling wine preferences by data mining from physicochemical properties.  
_Decision Support Systems_. __2009__, _47_, 547-553.

[Elsevier](http://dx.doi.org/10.1016/j.dss.2009.05.016)  
[Pre-press (pdf)](http://www3.dsi.uminho.pt/pcortez/winequality09.pdf)  
[bib](http://www3.dsi.uminho.pt/pcortez/dss09.bib)

__Question guiding investigation:__ Which chemical properties influence the 
quality of white wines?

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

library(ggplot2)
library(gridExtra)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load the Data
white.wine.df <- read.csv("wineQualityWhites.csv")
```

# Univariate Plots Section

__Preliminary information on the data set:__  
```{r echo=FALSE, message=FALSE, warning=FALSE}
str(white.wine.df)
summary(white.wine.df)
```

This data includes 4898 observations with 13 variables.

__Some histograms of organic acid contents and overall acidity (pH):__  
```{r echo=FALSE, message=FALSE, warning=FALSE}
ta.hist <- ggplot(data = white.wine.df, aes(x = fixed.acidity)) +
  geom_histogram() +
  labs(x = "Fixed acidity (g/dm^3 of tartaric acid)")

ca.hist <- ggplot(data = white.wine.df, aes(x = citric.acid)) +
  geom_histogram() +
  labs(x = "Citric acid content (g/dm^3)")

aa.hist <- ggplot(data = white.wine.df, aes(x = volatile.acidity)) +
  geom_histogram() +
  labs(x = "Volatile Acidity (g/dm^3 of acetic acid)")

ph.hist <- ggplot(data = white.wine.df, aes(x = pH)) +
  geom_histogram() +
  labs(x = "pH")

grid.arrange(ta.hist, ca.hist, aa.hist, ph.hist, ncol =2)
```

The various acid measures all have a normal distribution, some with a few
outliers.

__Histograms of sulfur measures:__  
```{r Sulfur measures, echo=FALSE, message=FALSE, warning=FALSE}
sulf.hist <- ggplot(data = white.wine.df, aes(x = sulphates)) +
  geom_histogram() +
  labs(x = "Sulphates (added potassium sulphate - g/dm^3)")

free.sulf.diox <- ggplot(data = white.wine.df, aes(x = free.sulfur.dioxide)) +
  geom_histogram() +
  labs(x = "free sulfur dioxide (mg / dm^3)")

total.sulf.diox <- ggplot(data = white.wine.df, aes(x = total.sulfur.dioxide)) +
  geom_histogram() +
  labs(x = "free sulfur dioxide (mg / dm^3)")

grid.arrange(sulf.hist, free.sulf.diox, total.sulf.diox, ncol = 2)
```

The sulfur measures also have normal distribution, though an extreme outlier
exists for the free sulfur dioxide histogram.

__Histograms of other wine attributes:__  
```{r echo=FALSE, message=FALSE, warning=FALSE}
res.sugar.hist <- ggplot(data = white.wine.df, aes(x = residual.sugar)) +
  geom_histogram() +
  labs(x = "Residual sugars (g/dm^3)")

density.hist <- ggplot(data = white.wine.df, aes(x = density)) +
  geom_histogram() +
  labs(x = "density (g/cm^3)")

grid.arrange(res.sugar.hist, density.hist, ncol = 2)
```

We can see an outlier in both residual sugars and density. My suspicion is
that they are the same wine (or few wines). Dissolved sugar increases the
density of water (which is 1.00 g/cm^3).

```{r, echo=FALSE, message=FALSE, warning=FALSE}
print(paste("High density wine:", 
            white.wine.df$density[white.wine.df$density > 1.02]))

print(paste("High residual sugar wine:", 
            white.wine.df$residual.sugar[white.wine.df$residual.sugar > 60]))

subset(white.wine.df, white.wine.df$density > 1.02)
```

We can see that they are indeed the same wine.

__Other quality measures:__  
Sodium chloride content  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
nacl.hist <- ggplot(data = white.wine.df, aes(x = chlorides)) +
  geom_histogram() +
  labs(x = "Sodium chloride (g/dm^3)")

summary(white.wine.df$chlorides)
```

There is a little bit of a tail in the sodium chloride data; most of the 
values are less than 0.05, yet there are a handful of values greater than 0.2 
Let's do a log base 10 transform on the x-axis.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
nacl.log10.hist <- ggplot(data = white.wine.df, aes(x = chlorides)) +
  geom_histogram() +
  labs(x = "Log 10 Sodium Chloride") +
  scale_x_log10()

grid.arrange(nacl.hist, nacl.log10.hist, ncol = 1)
```

We see that there really aren't many data points beyond 0.1 g/dm^3. This is
consistent with the mean not being far from the median.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = alcohol)) +
  geom_histogram(binwidth = 0.25) +
  labs(x = "Alcohol (ABV - % by volume)")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = white.wine.df, aes(x = quality)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Quality Rating (0 - 10)") +
  scale_x_continuous(limits = c(2, 10), breaks = seq(0, 10, 1))
```

The quality ratings are uniformly distributed around 6 (the mode).

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!